EDA

### 2024-11-29 other/EDA/EDA.qmd
fixed up top-buyers chart

df_filtered$year <- format(df_filtered$award_date, "%Y")

@@

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending, fill = buyer_category)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  scale_fill_manual(
    values = c(
      "National Defence" = "steelblue",
      "Public Services and Procurement Canada" = "orange",
      "Shared Services Canada" = "green",
      "Public Health Agency of Canada" = "purple",
      "Fisheries and Oceans Canada" = "red",
      "Other" = "gray"
    )
  ) +
  labs(
    title = "Spending by Selected and Other Buyers Per Year",
    x = "Year",
    y = "Total Spending",
    fill = "Buyer"
  ) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

>>>>

```{r}
library(ggplot2)
library(dplyr)
library(scales) # For formatting y-axis labels

library(ggplot2)
library(dplyr)
library(scales) # For formatting y-axis labels

# Filter out records from 2019
df_filtered <- df %>%
  filter(format(award_date, "%Y") != "2019")

# Extract the year from award_date
df_filtered$year <- format(df_filtered$award_date, "%Y")

@@

# Scale the total_spending to billions
grouped_df <- grouped_df %>%
  mutate(total_spending_billion = total_spending / 1e9)

# Create the grouped bar chart
ggplot(data = grouped_df, aes(x = year, y = total_spending_billion, fill = buyer_category)) +
  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.9), # Adjust dodge width
    width = 0.8 # Adjust bar width
  ) +
  scale_fill_manual(
    values = c(
      "National Defence" = "steelblue",
      "Public Services and Procurement Canada" = "orange",
      "Shared Services Canada" = "green",
      "Public Health Agency of Canada" = "purple",
      "Fisheries and Oceans Canada" = "red",
      "Other" = "gray"
    )
  ) +
  scale_y_continuous(
    labels = comma_format(accuracy = 1), # Formats y-axis labels without decimal places
    breaks = pretty_breaks(n = 5), # Ensures approximately 5 breaks like 5, 10, 15
    expand = expansion(mult = c(0, 0.05)) # Adds a small space above the highest bar
  ) +
  labs(
    title = "Spending by Selected and Other Buyers Per Year",
    subtitle = "Amounts are in Billions of CAD",
    x = "Year",
    y = "Total Spending (Billions)",
    fill = "Buyer",
    caption = "."
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1), # Rotate x-axis labels for readability
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.title = element_text(size = 14),
    legend.text = element_text(size = 12),
    plot.caption = element_text(hjust = 0.5, size = 10, face = "italic") # Centered caption
  )
